---
title: "Visualization 1"
output: github_document
date: "2025-09-30"
---

```{r, include = FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

Import weather data

```{r}
data("weather_df")
```


Making first plot:

```{r}

ggplot(data = weather_df, mapping = aes(x = tmin, y = tmax)) + geom_point()

```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Sitting in environment

```{r}
ggp_weather_scatterplot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```


Check that some rows are missing

```{r}
weather_df |> 
  filter(is.na(tmax))
```


## Fancier scatterplots

```{r}
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)
```

Where you find aesthetics matter

```{r}
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + geom_point(aes(color = name, size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE)
```

Use faceting real quick

```{r}
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + geom_point(aes(color = name), alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)

  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + geom_point(aes(color = name), alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE) + 
  facet_grid(name ~ .)

```

Let's make a more interesting scatterplot

```{r}
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + 
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

Learning Assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth)

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    t_max_fahr = tmax*(9/5) + 32,
    t_min_fahr = tmin*(9/5) + 32
  ) |> 
  ggplot(aes(x = t_min_fahr, y = t_max_fahr)) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")
```

## Small things

```{r}
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + 
#  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) 

  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + 
  geom_smooth(se = FALSE) +
  geom_point() 
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax))+
  geom_hex()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y =tmax))+
  geom_point(color = "purple")
```

## Univariate plots

```{r}
weather_df |> 
  ggplot(aes(x = tmin)) +
  geom_histogram(color = "white", fill = "red")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, color = name))+
  geom_histogram()

weather_df |> 
  ggplot(aes(x = tmin, fill = name))+
  geom_histogram()
```

how would I fix this? maybe facet?

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() + 
  facet_grid(name ~ .)
```

maybe a density plot?

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.2)
```

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot(aes(fill = name))
```

violin plots

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()
```

ridge plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = name, fill = name)) +
  geom_density_ridges()
```

Learning Assessment: Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable.


```{r}
weather_df |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density()

weather_df |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_histogram() + 
  facet_grid(name ~ .)

weather_df |> 
  filter(prcp > 5, prcp < 1000) |> 
  ggplot(aes(x = prcp, fill = name))+
  geom_density(alpha = 0.2)
```

## Save plots

```{r}
ggp_weather = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmin, fill = name))+
  geom_violin()

ggp_weather

ggsave("violin_plots.pdf", ggp_weather,
       width = 8, height = 6)

```

